Interpretable Neural Networks using EAGGA

### 2025-03-15 paper/main.tex
future outlook

\subsubsection{A-posteriori}
\label{sec-moo-post}
The disadvantage of a-priori methods is the missing knowledge of the interplay between a hyperparameter configuration and the trained model's
performance across objectives:
we can either restrict the search space to enforce some will, such as only using a maximum of 50\% of features, optimize for performance and use the
resulting model, or leave the search space unrestricted, adjust the loss function to incorporate all objectives and take its optimum.
In neither case do we know the impact our a-priori trade-off has on the final performance.
In practical applications, the impact on performance across all objectives is what matters.
This is the main advantage of a-posteriori methods: instead of implicitly defining a trade-off prior to training, we still evaluate multiple HP configurations
but keep the set of non-dominated solutions instead of just one that satisfies the trade-off criterion.
This is beneficial as it makes the relationship visible, a practicioner can now see, for instance, that a slight decrease in one objective might translate to
a significant improvement in another that would have been missed if the problem was optimized with a-priori methods.
\\
Aside from the usual baselines grid and random search there are multi-objetive Bayesian Optimization adaptations, mainly using one of two approaches:
\begin{enumerate}[label*=\arabic*.]
  \item Fitting a single surrogate model on scalarized objectives. \citet[pp. 54-56]{ParEGO} propose \textit{ParEGO}, which employs the augemented Tchebycheff function as
        scalarization to ensure the Pareto front is explored sufficiently.
  \item Fitting one surrogate per objective, then either 
  \begin{enumerate}[label*=\arabic*.]
    \item using one acquisition function per surrogate to return a set of promising candidate query points or
    \item using one overall acquisition function to aggregate the surrogates, such as \textit{EHI} as proposed by \citet[pp. 8f]{EHI}, which maximises
          the expected improvement of hypervolume.
  \end{enumerate}
\end{enumerate}

@@

\section{Conclusion and Future Outlook}
% Conclusion
The EAGGA framework has been shown to be able to preserve performance while increasing interpretabiltiy of XGBoost models on tabular data, which has long been 
a challenging discipline for neural networks. Following recent research suggesting that regularisation can improve
performance \citep[8]{NEURIPS2021_c902b497}, we extended EAGGA \citep{EAGGA} to neural networks, utilising its regularising properties to achieve similar
performance of neural networks as performance achieved with XGBoost and simultaneously increase interpretability of this notoriously uninterpretable model class.
As a consequence, we propose a new architecture allowing to model the equivalence relations of the EAGGA framework.
We create a model whose interpretation is active, attributional, and global in the framework of \citet[chap. 2]{survey_NN_interpretability}, and find
overall performance, as measured on the dominated hypervolume of (AUC, NF, NI, NNM) to be comparable to that of XGBoost fitted usign EAGGA.

(Future Outlook)
- due to high complexity of EAs, a MOBO implementation on augmented space (cf section \ref{sec-eagga}) would be very nice
- theoretical idea: use surrogate model with prior computed based on feasible combinations of equivalence classes
  * high value, 0 uncertainty around combinations that make no sense
    - e.g. those with a feature in multiple groups
    - e.g. those with two features in one group, one having monot incr, one monot decr
  * high uncertainty around combinations that would make sense to encourage exploring these regions as with GPs
  * this would likely prove to be computationally very expensive for higher numbers of features
- as more practical alternative: unrestricted MO BO on extended search space, cf section \ref{sec-eagga}, then acquisiton function taking into
  account the constraints, e.g. as presented by \citep{10.5555/3020751.3020778}, \citep{perrone2019constrainedbayesianoptimizationmaxvalue}

>>>>

\subsubsection{A-posteriori}
\label{sec-moo-post}
The disadvantage of a-priori methods is the missing knowledge of the interplay between a hyperparameter configuration and the trained model's
performance across objectives:
we can either restrict the search space to enforce some will, such as only using a maximum of 50\% of features, optimize for performance and use the
resulting model, or leave the search space unrestricted, adjust the loss function to incorporate all objectives and take its optimum.
In neither case do we know the impact our a-priori trade-off has on the final performance.
In practical applications, the impact on performance across all objectives is what matters.
This is the main advantage of a-posteriori methods: instead of implicitly defining a trade-off prior to training, we still evaluate multiple HP configurations
but keep the set of non-dominated solutions instead of just one that satisfies the trade-off criterion.
This is beneficial as it makes the relationship visible, a practicioner can now see, for instance, that a slight decrease in one objective might translate to
a significant improvement in another that would have been missed if the problem was optimized with a-priori methods.
\\
Aside from the usual baselines grid and random search there are multi-objetive Bayesian Optimization adaptations, mainly using one of two approaches:
\begin{enumerate}[label*=\arabic*.]
  \item Fitting a single surrogate model on scalarized objectives. \citet[pp. 54-56]{ParEGO} propose \textit{ParEGO}, which employs the augmented Tchebycheff function as
        scalarization to ensure the Pareto front is explored sufficiently.
  \item Fitting one surrogate per objective, then either 
  \begin{enumerate}[label*=\arabic*.]
    \item using one acquisition function per surrogate to return a set of promising candidate query points or
    \item using one overall acquisition function to aggregate the surrogates, such as \textit{EHI} as proposed by \citet[pp. 8f]{EHI}, which maximises
          the expected improvement of hypervolume.
  \end{enumerate}
\end{enumerate}

@@

\section{Conclusion and Future Outlook}
% Conclusion
The EAGGA framework has been shown to be able to preserve performance while increasing interpretabiltiy of XGBoost models on tabular data, which has long been 
a challenging discipline for neural networks. Following recent research suggesting that regularisation can improve NN performance on
tabular data \citep[8]{NEURIPS2021_c902b497}, we extend EAGGA \citep{EAGGA} to neural networks, utilising its regularising properties to achieve similar
performance with neural networks as with XGBoost and simultaneously increase interpretability of this notoriously uninterpretable model class.
As a consequence, we propose a new architecture allowing to model the equivalence relations of the EAGGA framework.
We create a model whose interpretation is active, attributional, and global in the framework of \citet[chap. 2]{survey_NN_interpretability}, and find
overall performance, as measured on the dominated hypervolume of (AUC, NF, NI, NNM) to be comparable to that of XGBoost fitted using EAGGA.
\\

% Future Outlook
Due to the low sample-efficiency of evolutionary algorithms, a multi-objective Bayesian Optimization variant on the augmented search space
(refer Section \ref{sec-eagga}) could deliver the strongest training time improvement.
There are two avenues to explore this, either via a change to the surrogate model or by adapting the acquisition function.
A surrogate model using a custom computed prior with high values and no uncertainty in unfeasible regions and regular uncertainty in feasible regions could be a solution.
An unfeasible region is for instance one, where a feature is in multiple equivalence classes, or where two features in the same equivalence class have opposing
monotonicity constraints.
The second option, adapting the acquisition function, may prove to be computationally less complex for datasets with many features.
This would mean using an unrestricted surrogate model and then taking account the constraints in the acquisition function. Similar approaches have been
tried on unrelated problems by \citep{10.5555/3020751.3020778} and \citep{perrone2019constrainedbayesianoptimizationmaxvalue}.
